Distributed Algorithms for Sparse Matrix-Vector Multiplication

### 2015-04-27 presentation.Rmd
More updates for pres

@@

@@

@@

@@

- A warp consisting of w threads (or rows)
- A list of cache line mappings $C_i$, for each thread $t_i$, corresponding to k memory elements
- A wxk scheduling table S where each row i can be any permutation of the $C_i$.

The object is to minimize $z(S) = \sum_{j=0}^{k-1}|\cap_{i=0}^{w-1} S_{i,j}|$.

## Cache Heuristic
- Greedy approach where the solution is incremenetally constructed from row 0 to row w - 1.
- Each permutation is selected in an attempt to maximize overlapping with rows already placed.
- $O(k)$ steps to find overlapping position for a single warp.
- Total time complexity of $O(k^2n)$
- Since k is typically very small relative to n due to the sparse nature of the matrices involved, the time complexity is approximately $O(n)$.

@@

## Conclusion
- Incremental speedup
-

@@

@@

## Conclusion

>>>>

```{r, results='asis', echo=FALSE}
library(xtable)

set.seed(20150427)

print(xtable(matrix(rbinom(40, 1, .05), ncol = 5)), type = "html", comment = FALSE)
```

@@

```{r, results='asis', echo=FALSE}
set.seed(20150427)

print(xtable(matrix(rbinom(40, 1, .05), ncol = 5)), type = "html", comment = FALSE)
```

@@

## GPU Revolution
- Large pool of threads
- Processing units called Streaming Multiprocessors (SMs)
- Each SM contains computing cores called Streaming Processors (SPs)
- Example: NVIDIA GTX 580 contains 16 SMs, each with 32 SPs, for a potential total of 512 operations per clock cycle!
- However, threads within the same warp (group of 32 threads) must execute the same instructions, or there is a performance hit.

@@

## Example Cache Request
<img src="images/Warp.png" width="500px">

## Example Cache Request (Continued)
<img src="images/Perm.png" width="600px">

@@

- A warp consisting of w threads (or rows)
- A list of cache line mappings $C_i$, for each thread $t_i$, corresponding to k memory elements
- A $w \times k$ scheduling table S where each row i can be any permutation of the $C_i$.

The object is to minimize $z(S) = \sum_{j=0}^{k-1}|\cap_{i=0}^{w-1} S_{i,j}|$. This is the **sum of cache line intersections along the vertical direction in the scheduling table S**.

## Cache Heuristic
- Greedy approach where the solution is incrementally constructed from row 0 to row w - 1.
- Each permutation is selected in an attempt to maximize overlapping with rows already placed.
- $O(k)$ steps to find overlapping position for a single warp.
- Total time complexity of $O(k^2n)$
- Since k is typically very small relative to n due to the sparse nature of the matrices involved, the time complexity is approximately $O(n)$.

@@

## Conclusion
- Incremental speedup
- A little bit outside the scope of our class
- Can we find a more distributed algorithm? (Rather than thread level...)

@@

## Distributed Memory Parallel Computers (DMPC)
- Each processor has local memory (no global shared memory)
- Communication via message passing
- Processors take either communication or computation steps
- A computation step can be "idle"
- Time complexity of a DMPC is the total number of communication and computation steps.

@@

## Conclusion
- Speedup linear in the number of processors (optimal)
- Lots of details and proofs available in paper
